---
title: "FlowSOM_CVID"
output: html_document
editor_options: 
  chunk_output_type: inline
  
---



## Set working directory for entire Rmd
```{r setup}
knitr::opts_knit$set(root.dir = "../")
```

## Loading in packages
```{r}
suppressPackageStartupMessages(library(tidyverse))

suppressPackageStartupMessages(library(flowCore))
suppressPackageStartupMessages(library(flowWorkspace))
suppressPackageStartupMessages(library(FlowSOM))
suppressPackageStartupMessages(library(MEM))
source("R_Scripts/FlowSOM_functions.R")
seed <- 1

```

## On what tube will you be working on? 

```{r}
tube <- 5

if (tube == 1){
  # PBMCs
  markers_to_cluster_idc <- c(1,4,7:16,18)
} else if (tube == 2) {
  # B cells
  markers_to_cluster_idc <- c(7:13,16,18)
} else if (tube == 5) {
  
# T cells
  markers_to_cluster_idc <- c(7:13,16,18)
} 
```


## Do you include the other PID?
```{r}
HC_CVID <- c("HC", "CVID")
PID_HC <- c(HC_CVID, "IgG_deficiency", "IgG_subclass_deficiency", "IgM_deficiency",
            "IgA_deficiency")

group_to_use <- PID_HC  # Choose between HC_CVID and PID_HC
PID <- "_PID_New_preprocessing" #Set to empty character if you are only working with HC en CVID (otherwise: _PID)
```


## List correct direcories
```{r}
data_dir = paste0("FlowSOM/Tube",tube,"/Data",PID)
fcs_dir = paste0("Preprocessing/Tube",tube)
figure_dir <- paste0("FlowSOM/Tube",tube,"/Figures", PID)
```

# Make directories if they don't exist yet
```{r}
suppressWarnings(dir.create(data_dir))
suppressWarnings(dir.create(figure_dir))
```


# FlowSOM analysis Human blood samples CVID

## Overview of the meta_data

```{r}
meta_data_file <- "../../PhD/CVID/Clinical_features_data.txt"
meta_data <- read.delim(meta_data_file, header = TRUE, sep = "\t", 
                        check.names = FALSE) %>% 
  dplyr::filter(diagnosis_group %in% group_to_use) %>%
  dplyr::filter(SAMPLE_ID != "PIDHC081")

if(length(group_to_use) == length(PID_HC)){
  meta_data$diagnosis_group <- as.character(meta_data$diagnosis_group)
  meta_data[!(meta_data$diagnosis_group %in% HC_CVID),"diagnosis_group"] <- "PID"
  meta_data$diagnosis_group <- as.factor(meta_data$diagnosis_group)
}

# PIDHC081 is removed because he has too few cells for tube 2
rownames(meta_data) <- meta_data$SAMPLE_ID

meta_data
```

## List files and select only HC and CVID files and/or other PID files

```{r}
allfiles <- list.files(fcs_dir, pattern = "^P.*_pp.fcs$", recursive = TRUE)
files <- allfiles[grep(paste(rownames(meta_data), collapse="|"), allfiles)]
```

## Make an aggregate file for all HC and CVID patients and/or other PID patiens 
# Included quality check

```{r}

set.seed(seed)
suppressWarnings(ff_agg <- AggregateFlowFrames(file.path(fcs_dir,files), 
                              cTotal = 3000000,
                              writeOutput  = TRUE,
                              outputFile = file.path(data_dir,
                                                     paste0("aggregate",PID,
                                                            seed,".fcs"))))
table(ff_agg@exprs[,"File"])

markers <- ff_agg@parameters@data[,"desc"]
markers[is.na(markers)] <- ff_agg@parameters@data[,"name"][is.na(markers)]
markers <- gsub("#.*", "", markers)
names(markers) <- colnames(ff_agg)
  
# Define the markers you want to use to identify the cellTypes
# Markers after the start gate
markers_to_cluster <-  colnames(ff_agg)[markers_to_cluster_idc]
# Define the markers you want to plot
markers_to_plot <- colnames(ff_agg)[markers_to_cluster_idc]

# Quality Control: all files should behave similarly as we will use them
# aggregated together. High populations should be at approximately the same MFI!
# We check this for all the markers we might be interested in at some point.

PlotMarkerMFI(files = files,
              markers_to_plot = markers_to_plot,
              marker_names = markers,
              ff = ff_agg,
              figure_dir = figure_dir)
  

```


## Set number of meta-clusters and clusters for the different panels
```{r}
if(tube == 1){
  nclus <- 10
  nmetaclus <- 14
} else if(tube == 2){
  nclus <- 10
  nmetaclus <- 18
} else if(tube == 5){
  nclus <- 10
  nmetaclus <- 18
} 
```


## Make the FlowSOM object
```{r}
ff <- read.FCS(file.path(data_dir, paste0("aggregate",PID, seed,".fcs")))
ff_agg <- ff
ff <- ff[,-c(20,21)]

markers <- ff@parameters@data[,"desc"]
markers[is.na(markers)] <- ff@parameters@data[,"name"][is.na(markers)]
markers <- gsub("#.*", "", markers)

names(markers) <- colnames(ff)


markers

fsom <- FlowSOM(ff,
                colsToUse = markers_to_cluster_idc,
                xdim = nclus, ydim = nclus,
                nClus = nmetaclus,
                seed = seed)

saveRDS(fsom,
     file = file.path(data_dir,
                      paste0("FlowSOM",
                             PID, "_",
                             nclus, "_", 
                             nmetaclus, "_",
                             seed,".RDS")))

```

# Plot FlowSOM tree with labeling of clusters and meta-clusters
```{r}
library(RColorBrewer)

fsom <- readRDS( file = file.path(data_dir, 
                                  paste0("FlowSOM",
                                         PID, "_",
                                         nclus, "_", 
                                         nmetaclus, "_",
                                         seed,".RDS")))

set.seed(seed)

fsom$FlowSOM$prettyColnames <- sub(" .*", "", fsom$FlowSOM$prettyColnames)


pdf(file.path(figure_dir,
              paste0("FlowSOM", PID, "_",
                             nclus, "_", 
                             nmetaclus, "_",
                             seed,"_stars.pdf")),
    useDingbats = FALSE)
PlotStars(UpdateNodeSize(fsom$FlowSOM,maxNodeSize = 10, reset = TRUE),
          backgroundValues = fsom$metaclustering, backgroundColor =  colorRampPalette(brewer.pal(12, "Set3"))(nmetaclus))
PlotNumbers(UpdateNodeSize(fsom$FlowSOM, maxNodeSize = 10, reset=TRUE))
PlotLabels(UpdateNodeSize(fsom$FlowSOM, maxNodeSize = 10, reset = TRUE), labels = fsom$metaclustering)
dev.off()
```


# Adapted plotting with different scaling (see FlowSOM_functions.R)
```{r}
library(RColorBrewer)


fsom <- readRDS( file = file.path(data_dir, 
                                  paste0("FlowSOM",
                                         PID, "_",
                                         nclus, "_", 
                                         nmetaclus, "_",
                                         seed,".RDS")))

fsom_original <- fsom
fsom <- UpdateNodeSize(fsom$FlowSOM,maxNodeSize = 10, reset = TRUE)

set.seed(seed)
fsom$FlowSOM$prettyColnames <- sub(" .*", "", fsom$FlowSOM$prettyColnames)


PlotStars_adapted(UpdateNodeSize(fsom$FlowSOM,maxNodeSize = 10, reset = TRUE),
          backgroundValues = fsom$metaclustering, backgroundColor =  colorRampPalette(brewer.pal(12, "Set3"))(nmetaclus), scaling = "MaxValue")
PlotNumbers(UpdateNodeSize(fsom$FlowSOM, maxNodeSize = 10, reset=TRUE))
PlotLabels(UpdateNodeSize(fsom$FlowSOM, maxNodeSize = 10, reset = TRUE), labels = fsom$metaclustering)


```

# Map all files on full tree to calculate purity and to plot the manual labels on the tree
```{r}
fsom <- readRDS( file = file.path(data_dir, 
                                  paste0("FlowSOM",
                                         PID, "_",
                                         nclus, "_", 
                                         nmetaclus, "_",
                                         seed,".RDS")))


pdf(file.path(figure_dir,
              paste0("FlowSOM", PID, "_",
                             nclus, "_", 
                             nmetaclus, "_",
                             seed,"_individual_files.pdf")),
    useDingbats = FALSE)

PlotStars(UpdateNodeSize(fsom$FlowSOM,maxNodeSize = 10, reset = TRUE),
          backgroundValues = fsom$metaclustering)


manual_gated_cells <- c()
metacluster_cells <- c()
cluster_cells <- c()

patient <- meta_data$SAMPLE_ID[80]


for(patient in meta_data$SAMPLE_ID){
  resting_samples <- (length(meta_data$SAMPLE_ID) - which(meta_data$SAMPLE_ID == patient)) + 1
  print(paste0("Nog ", resting_samples, " files te berekenen."))
    file <- grep(patient, files,value = TRUE)
  if (length(file) > 1){
    file <- file[length(file)]
  }
  if (length(file) == 0){
    next
  }
  ff <- read.FCS(file.path(fcs_dir,file))
  # Load the corresponding manual gating
  load(file.path(fcs_dir,
                 gsub("_pp.fcs","_manual.Rdata",file)))
  
  # Map the new file to the flowSOM structure computed before
  fsom_tmp <- NewData(fsom$FlowSOM,ff)
  
  if(tube == 5){
    manual$manual <- factor(manual$manual,levels(manual$manual)[c(13,1:12)])
  }
  
  PlotStars(fsom_tmp, 
            main=patient)
  PlotPies(UpdateNodeSize(fsom_tmp,reset = TRUE, maxNodeSize = 10),
            cellTypes = manual$manual, 
           main=patient)

  manual_gated_cells <- c(manual_gated_cells, as.character(manual$manual))
  cluster_cells <- c(cluster_cells, GetClusters(fsom_tmp))
  metacluster_cells <- c(metacluster_cells, GetMetaclusters(fsom_tmp, fsom$metaclustering))
  
  
      
  }
dev.off()

purity_clusters <- Purity(manual_gated_cells, cluster_cells, weighted = F)
purity_metaclusters <- Purity(manual_gated_cells, metacluster_cells, weighted = F)

purity_clusters
purity_metaclusters
FMeasure(manual_gated_cells, metacluster_cells)



```


# Plot the 2d plots for the different populations to see how well the metaclusters predict the populations
```{r}

ff <- read.FCS(file.path(fcs_dir, files[71]))
#flowSOM.res <- FlowSOM(ff, colsToUse = markers_to_cluster_idc, nClus = nmetaclus, xdim = 10, ydim = 10 )

fsom <- readRDS( file = file.path(data_dir, 
                                  paste0("FlowSOM",
                                         PID, "_",
                                         nclus, "_", 
                                         nmetaclus, "_",
                                         seed,".RDS")))

pdf(paste0(figure_dir,"/Markeroverview_origTree.pdf"),
    width =  14, height = 10, useDingbats = FALSE)

fsom$FlowSOM$prettyColnames <- sub(" .*", "", fsom$FlowSOM$prettyColnames)


PlotStars_adapted(UpdateNodeSize(fsom$FlowSOM,maxNodeSize = 10, reset = TRUE),
          backgroundValues = fsom$metaclustering, backgroundColor = colorRampPalette(brewer.pal(12, "Set3"))(nmetaclus), scaling = "MaxValue")
PlotNumbers(UpdateNodeSize(fsom$FlowSOM,reset=TRUE, maxNodeSize = 10))
PlotLabels(UpdateNodeSize(fsom$FlowSOM, maxNodeSize = 10, reset = TRUE), labels = fsom$metaclustering)

if (tube == 2){
  PlotStars_adapted(UpdateNodeSize(fsom$FlowSOM, maxNodeSize = 10, reset = TRUE),
            markers = get_channels(ff, c("IgA", "IgD", "IgM","IgG")), 
            backgroundValues = fsom$metaclustering, backgroundColor =  colorRampPalette(brewer.pal(12, "Set3"))(nmetaclus),  scaling = "MaxValue")
  PlotStars_adapted(UpdateNodeSize(fsom$FlowSOM, maxNodeSize = 10, reset = TRUE),
            markers = get_channels(ff, c("CD38", "CD138", "CD21", "CD24", "CD27")),  backgroundValues = fsom$metaclustering, backgroundColor =  colorRampPalette(brewer.pal(12, "Set3"))(nmetaclus), scaling = "MaxValue")
}

set.seed(1)
subset <- sample(seq_len(nrow(fsom$FlowSOM$data)),
                 100000)
subset <- FlowSOMSubset(fsom$FlowSOM, subset)

for(i in seq(1, nmetaclus, by = 3)){

  
  if (tube == 1){
    
    marker_list <- list(c("PE-CF594", "SSC-A"),
                        c("APC", "AF700"),
                        c("PerCP-Cy5-5", "FITC"),
                        c("PE-Cy7", "PerCP-Cy5-5"),
                        c("Pacific Blue", "PE"),
                        c("PE", "PE-Cy5"))
    
  } else if (tube == 2){
  
    marker_list <- list(c("PE-Cy7", "PE-CF594"),
                        c("PE-Cy7", "PerCP-Cy5-5"),
                        c("PE-Cy5", "BV605"),
                        c("APC", "PE"),
                        c("FITC", "BV605"))
  } else if (tube == 3){
      marker_list <- list(c( "Pacific Blue-A", "APC-A"),
                          c("FITC", "FSC-A"),
                          c("APC-Cy7", "FSC-A"),
                          c("Qdot 605", "FSC-A"))
  }
  else if (tube == 5){
      marker_list <- list(c("Pe-Cy7", "AF700"),
                          c("Pacific Blue", "FITC"),
                          c("PE-Cy5", "PE"),
                          c("PE-Cy5", "PE-CF594"),
                          c("PE-Cy5","V605"),
                          c("PE", "PerCP-Cy5-5"))
    }
  
  PlotOverview2D(list(FlowSOM = subset, metaclustering = fsom$metaclustering),
                 marker_list,
                 metaclusters = c(i:(i+2)),
                 ff = ff)
}

dev.off()



```


# Make densityplots per metacluster (See density_functionplot.R)
```{r}
pdf(paste0(figure_dir,"/Densityoverview_OnePatient.pdf"),
    width =  14, height = 10)

source("R_Scripts/density_functionplot.R")

  ff <- read.FCS(file.path(fcs_dir, files[71]))
  fsom_tmp <- NewData(fsom$FlowSOM,ff)


for (i in c(1,8)){
suppressWarnings(PlotClusters1D(list(FlowSOM = fsom_tmp, metaclustering = fsom$metaclustering ), markerlist = names(fsom$FlowSOM$prettyColnames)[markers_to_cluster_idc], metaclusters = c(i:(i+6)), textsize = 5))
}
dev.off()

```

















