---
title: "FlowSOM_PRISM"
output: html_document
---


## Set working directory for entire Rmd
```{r setup}
knitr::opts_knit$set(root.dir = "../")
```

# Load in packages
```{r}
suppressPackageStartupMessages(library(tidyverse))

suppressPackageStartupMessages(library(flowCore))
suppressPackageStartupMessages(library(flowWorkspace))
suppressPackageStartupMessages(library(FlowSOM))
suppressPackageStartupMessages(library(MEM))
suppressPackageStartupMessages(library(qsub))

seed <- 1

```

## On what tube will you be working on? 

```{r}
tube <- 5

if (tube == 1){ #PBMCs
  markers_to_cluster_idc <- c(1,4,7:16,18)
} else if (tube == 2) { # B cells
  markers_to_cluster_idc <- c(7:13,16,18)
}  else if (tube == 5) { # T cells
  markers_to_cluster_idc <- c(7:13,16,18)
}
```


## Do you include the other PID?
```{r}
HC_CVID <- c("HC", "CVID")
PID_HC <- c(HC_CVID, "IgG_deficiency", "IgG_subclass_deficiency", "IgM_deficiency",
            "IgA_deficiency")

# Use this one if you want to map all the files including the HFM on the trees as well
PID_HC <- c(HC_CVID, "IgG_deficiency", "IgG_subclass_deficiency", "IgM_deficiency",
            "IgA_deficiency", "HFM")


group_to_use <- PID_HC  # Choose between HC_CVID and PID_HC
PID <- "_PID_New_preprocessing" #Set to empty character if you are only working with HC en CVID (otherwise: _PID)

seed <- 1
```


## List correct direcories
```{r}
data_dir = paste0("FlowSOM/Tube",tube,"/Data",PID,"/Crossvalidation")
fcs_dir = paste0("Preprocessing/Tube",tube)
figure_dir <- paste0("FlowSOM/Tube",tube,"/Figures", PID, "/Crossvalidation")
aggregate_dir <- file.path(data_dir, "Aggregates")
fsom_dir <- file.path(data_dir, "fsom_trees")
fsom_objects <- file.path(data_dir, "fsom_objects")

```


# Make directories if they don't exist yet
```{r}
suppressWarnings(dir.create(data_dir))
suppressWarnings(dir.create(figure_dir))
suppressWarnings(dir.create(fsom_dir))
suppressWarnings(dir.create(aggregate_dir))
suppressWarnings(dir.create(fsom_objects))
```


# FlowSOM analysis Human blood samples CVID run on the cluster

## Overview of the meta_data

```{r}
meta_data_file <- "../../PhD/CVID/Clinical_features_data.txt"
meta_data <- read.delim(meta_data_file, header = TRUE, sep = "\t", 
                        check.names = FALSE) %>% 
  dplyr::filter(diagnosis_group %in% group_to_use) %>%
  dplyr::filter(SAMPLE_ID != "PIDHC081")

if(length(group_to_use) == length(PID_HC)){
  meta_data$diagnosis_group <- as.character(meta_data$diagnosis_group)
  meta_data[!(meta_data$diagnosis_group %in% c(HC_CVID, "HFM")),"diagnosis_group"] <- "PID"
  meta_data$diagnosis_group <- as.factor(meta_data$diagnosis_group)
}

# PIDHC081 is removed because he has too few cells for tube 2
rownames(meta_data) <- meta_data$SAMPLE_ID

meta_data
```


# Set up qsub config 
```{r}
qsub_config = create_qsub_config(
  remote = "",
  local_tmp_path = "C:/Users//Documents/.r2gridengine",
  remote_tmp_path = "/scratch/irc/personal//r2gridengine",
  modules = "R/x86_64/3.5.1",
  memory = "4G",
  wait = TRUE, 
  remove_tmp_folder = TRUE,
  name = "FlowSOM_exp13",
  max_wall_time = "500:00:00",
  stop_on_error = TRUE
)
qsub::set_default_qsub_config(qsub_config, permanent = T)

```

## List files and select only HC and CVID files and/or other PID files

```{r}
allfiles <- list.files(fcs_dir, pattern = "^P.*_pp.fcs$", recursive = TRUE)
files <- allfiles[grep(paste(rownames(meta_data), collapse="|"), allfiles)]
```

# Make FlowSOM model per experiment day --> save in RDS object


```{r}
if(tube == 1){
  nclus <- 10
  nmetaclus <- 14
} else if(tube == 2){
  nclus <- 10
  nmetaclus <- 18
}else if(tube == 5){
  nclus <- 10
  nmetaclus <- 18
} 
```



## Select correct files for the experiment day + make variables for patients #DONE!!!
```{r}

experiment_day <- 22

pdf(file = file.path(figure_dir,"fsom_stars_metaclusters.pdf"))

for (experiment_day in c(13:33)){
  
  print(paste0("Experiment day ",experiment_day))
  
  train_files <- files[-grep(paste0("Exp_",experiment_day), files)]
  train_files
  
  set.seed(seed)
  ## Make aggregate file for selected files
  suppressWarnings(ff_agg <- AggregateFlowFrames(file.path(fcs_dir,train_files), 
                                cTotal = 3000000,
                                writeOutput  = TRUE,
                                outputFile = file.path(aggregate_dir,
                                                       paste0("Without_Experiment_day_",
                                                              experiment_day, 
                                                              "_aggregate",
                                                              seed,".fcs"))))
  #table(ff_agg@exprs[,"File"])
  ff_agg <- read.FCS(file = file.path(aggregate_dir, paste0("Without_Experiment_day_",
                                                           experiment_day,
                                                           "_aggregate", seed, ".fcs")))
  
  ff_agg <- ff_agg[,-c(20,21)]
  
  markers_to_cluster <-  colnames(ff_agg)[markers_to_cluster_idc]
  
  
  ## Make fsom model
  fsom <- FlowSOM(ff_agg,
                  colsToUse = markers_to_cluster_idc,
                  xdim = nclus, ydim = nclus,
                  nClus = nmetaclus,
                  seed = seed)
  
  saveRDS(fsom, file = file.path(fsom_dir, 
                                 paste0("fsom_without_experiment_day_", 
                                        experiment_day, ".Rds")))
  #fsom <- readRDS(paste0("FlowSOM/Tube1/Data/Crossvalidation/fsom_trees/fsom_without_experiment_day_",experiment_day,".Rds"))
  PlotStars(UpdateNodeSize(fsom$FlowSOM,maxNodeSize = 10, reset = TRUE),
            backgroundValues = fsom$metaclustering, main = paste0("Experiment ", experiment_day, " was left out."))
  PlotNumbers(UpdateNodeSize(fsom$FlowSOM,reset=TRUE))
  PlotLabels(UpdateNodeSize(fsom$FlowSOM, maxNodeSize = 10, reset = TRUE), labels = fsom$metaclustering)
  
}


dev.off()


```



# Script for cross-validation on the cluster --> save all matrices in Rdata object
# Extraction of FlowSOM features per crossvaldation FlowSOM tree 
```{r}

experiment_day <- 13


  train_files <- files ## Make sure that you take all the files to map!!
  #train_files
  #f <- lapply(file.path(fcs_dir, train_files), read.FCS)
  fcs_dir <- paste0("/group/irc/personal/anneliese/Preprocessed_files/Tube", tube)
for (experiment_day in c(23:33)){
  print(paste0("Experiment day ", experiment_day))
  
  #Just for the entire flowSOM tree
  #fsom <- readRDS(file =  paste0("FlowSOM/tube", tube, "/Data", PID, "/FlowSOM", PID, "_",nclus,"_",nmetaclus,"_",seed,".RDS"))
  
  fsom <- readRDS(file = paste0(fsom_dir,"/fsom_without_experiment_day_",
                                experiment_day,
                                ".Rds"))

  handle <- qsub_lapply(
    X = seq_len(length(train_files)),
    qsub_config = override_qsub_config(name = "FlowSOM_Cross_Validation", wait = FALSE, memory = "5G"),
    qsub_environment = c("fsom", "markers_to_cluster_idc", "nclus", "nmetaclus", "fcs_dir", "experiment_day", "train_files"),
    FUN = function(x) {  
            library(dplyr)
            library(flowCore)
      f <- read.FCS(file.path(fcs_dir,train_files[x]))

      counts_t <- matrix(0, 
                     nrow = 1, 
                     ncol = nclus*nclus, 
                     dimnames = list("", 
                                     as.character(seq_len(nclus*nclus)))) 
    meta_clusters_t <- matrix(0, 
                              nrow = 1, 
                              ncol = nmetaclus, 
                              dimnames = list("", 
                                              as.character(seq_len(nmetaclus)))) 
  
    
    fsom_tmp <- FlowSOM::NewData(fsom$FlowSOM,f)
    cells_clusters <- FlowSOM::GetClusters(fsom_tmp) %>% table()
    counts_t[,names(cells_clusters)] <- cells_clusters
    cells_metaclusters <- fsom$metaclustering[fsom_tmp$map$mapping[,1]] %>% table()
    meta_clusters_t[,names(cells_metaclusters)] <- cells_metaclusters
    MFI_matrices_clusters <-  matrix(0,
                   nrow = dim(counts_t)[2],
                   ncol = length(markers_to_cluster_idc),
                   dimnames = list(as.character(seq_len(dim(counts_t)[2])), 
                                   names(fsom$FlowSOM$prettyColnames[markers_to_cluster_idc])))
    for (colour in colnames(MFI_matrices_clusters)){
      print(colour)
      MFI_matrices_clusters[, colour] <- 
        sapply(rownames(MFI_matrices_clusters), function(i) 
      median(f@exprs[FlowSOM::GetClusters(fsom_tmp) == i,colour]))
    }
    
    MFI_matrices_clusters[is.na(MFI_matrices_clusters)] <- 0    
    MFI_matrices_metaclusters <-  matrix(0,
                   nrow = dim(meta_clusters_t)[2],
                   ncol = length(markers_to_cluster_idc),
                   dimnames = list(as.character(seq_len(dim(meta_clusters_t)[2])), 
                                   names(fsom$FlowSOM$prettyColnames[markers_to_cluster_idc])))
    for (colour in colnames(MFI_matrices_metaclusters)){
      MFI_matrices_metaclusters[, colour] <- 
        sapply(rownames(MFI_matrices_metaclusters), function(i) 
      median(f@exprs[fsom$metaclustering[fsom_tmp$map$mapping[,1]] == i,colour]))
    }
    MFI_matrices_metaclusters[is.na(MFI_matrices_metaclusters)] <- 0
        
    colnames(MFI_matrices_clusters) <- 
      sub(" .*", "",fsom$FlowSOM$prettyColnames[markers_to_cluster_idc])
    colnames(MFI_matrices_metaclusters) <- 
      sub(" .*", "",fsom$FlowSOM$prettyColnames[markers_to_cluster_idc])
    
    percentages_cluster_to_meta_cluster <- 
      sapply(seq_len(dim(counts_t)[2]), 
             function(x)counts_t[x]/sum(counts_t[which(fsom$metaclustering == fsom$metaclustering[x])]))
        
    percentages_clusters <- 
      sapply(seq_len(dim(counts_t)[2]), 
             function(x)counts_t[x]/sum(counts_t))
        
    percentages_metaclusters <- 
      sapply(seq_len(dim(meta_clusters_t)[2]), 
             function(x)meta_clusters_t[x]/sum(meta_clusters_t))
        
    result <- list("Patient"= f@description$`$SRC`,
                   "counts_clusters" = counts_t, 
                   "counts_meta_clusters" = meta_clusters_t, 
                   "MFI_clusters" = MFI_matrices_clusters, 
                   "MFI_metaclusters" = MFI_matrices_metaclusters, 
                   "percentages_cluster2metacluster" = percentages_cluster_to_meta_cluster, 
                   "percentages_clusters" = percentages_clusters, 
                   "percentages_metaclusters" = percentages_metaclusters)
        })
  
  #Just for the entire flowSOM tree
  #saveRDS(handle, file = paste0("FlowSOM/tube", tube, "/Data", PID, "/fsom_objects.Rds"))

  saveRDS(handle, file = paste0(fsom_objects,"/fsom_objects_experiment_day_handle_cluster",experiment_day,".Rds"))
}

```



#Since we wait on the results of the cluster, here we can retrieve the results

```{r}
experiment_day <- 13
for (experiment_day in c(23:33)){
  experiment <- readRDS(file = paste0(fsom_objects, "/fsom_objects_experiment_day_handle_cluster", experiment_day, ".Rds"))
  handle <- qsub_retrieve(experiment)
  saveRDS(handle, file = paste0(fsom_objects, "/fsom_objects_experiment_day", experiment_day, ".Rds"))
}
```


# Without using the cluster
```{r}


experiment_day <- 13

for (experiment_day in c(13:33)){

  
  fsom <- readRDS(file = paste0(fsom_dir,"/fsom_without_experiment_day_",
                                experiment_day,
                                ".Rds"))
  #train_files <- files[grep("PIDHC085", files)]
  #train_files
  file <- files[17]
  final <- list()
  for (file in files){
    counts_t <- matrix(0, 
                     nrow = 1, 
                     ncol = nclus*nclus, 
                     dimnames = list("", 
                                     as.character(seq_len(nclus*nclus)))) 
    meta_clusters_t <- matrix(0, 
                              nrow = 1, 
                              ncol = nmetaclus, 
                              dimnames = list("", 
                                              as.character(seq_len(nmetaclus)))) 
  
    
    ff <- read.FCS(file.path(fcs_dir, file))
    fsom_tmp <- FlowSOM::NewData(fsom$FlowSOM,ff)
    cells_clusters <- FlowSOM::GetClusters(fsom_tmp) %>% table()
    counts_t[,names(cells_clusters)] <- cells_clusters
    cells_metaclusters <- fsom$metaclustering[fsom_tmp$map$mapping[,1]] %>% table()
    meta_clusters_t[,names(cells_metaclusters)] <- cells_metaclusters
    MFI_matrices_clusters <-  matrix(0,
                   nrow = dim(counts_t)[2],
                   ncol = length(markers_to_cluster_idc),
                   dimnames = list(as.character(seq_len(dim(counts_t)[2])), 
                                   names(fsom$FlowSOM$prettyColnames[markers_to_cluster_idc])))
    for (colour in colnames(MFI_matrices_clusters)){
      print(colour)
      MFI_matrices_clusters[, colour] <- 
        sapply(rownames(MFI_matrices_clusters), function(i) 
      median(ff@exprs[FlowSOM::GetClusters(fsom_tmp) == i,colour]))
    }
    
    MFI_matrices_clusters[is.na(MFI_matrices_clusters)] <- 0    
    MFI_matrices_metaclusters <-  matrix(0,
                   nrow = dim(meta_clusters_t)[2],
                   ncol = length(markers_to_cluster_idc),
                   dimnames = list(as.character(seq_len(dim(meta_clusters_t)[2])), 
                                   names(fsom$FlowSOM$prettyColnames[markers_to_cluster_idc])))
    for (colour in colnames(MFI_matrices_metaclusters)){
      MFI_matrices_metaclusters[, colour] <- 
        sapply(rownames(MFI_matrices_metaclusters), function(i) 
      median(ff@exprs[fsom$metaclustering[fsom_tmp$map$mapping[,1]] == i,colour]))
    }
    MFI_matrices_metaclusters[is.na(MFI_matrices_metaclusters)] <- 0
        
    colnames(MFI_matrices_clusters) <- 
      sub(" .*", "",fsom$FlowSOM$prettyColnames[markers_to_cluster_idc])
    colnames(MFI_matrices_metaclusters) <- 
      sub(" .*", "",fsom$FlowSOM$prettyColnames[markers_to_cluster_idc])
    
    percentages_cluster_to_meta_cluster <- 
      sapply(seq_len(dim(counts_t)[2]), 
             function(x)counts_t[x]/sum(counts_t[which(fsom$metaclustering == fsom$metaclustering[x])]))
        
    percentages_clusters <- 
      sapply(seq_len(dim(counts_t)[2]), 
             function(x)counts_t[x]/sum(counts_t))
        
    percentages_metaclusters <- 
      sapply(seq_len(dim(meta_clusters_t)[2]), 
             function(x)meta_clusters_t[x]/sum(meta_clusters_t))
        
    result <- list("Patient"= ff@description$`$SRC`,
                   "counts_clusters" = counts_t, 
                   "counts_meta_clusters" = meta_clusters_t, 
                   "MFI_clusters" = MFI_matrices_clusters, 
                   "MFI_metaclusters" = MFI_matrices_metaclusters, 
                   "percentages_cluster2metacluster" = percentages_cluster_to_meta_cluster, 
                   "percentages_clusters" = percentages_clusters, 
                   "percentages_metaclusters" = percentages_metaclusters)
    
    final[[ff@description$`$SRC`]] <- result
      }
      
  saveRDS(final, file = paste0(fsom_objects,"/fsom_objects_experiment_day",experiment_day,".Rds"))
}


```

