---
title: "Manual_Full_analysis"
output: html_document
---


## This document is to do an analaysis on a full Manual model in order to find relevant features to make tSNE plots and to inspect most important features


## Set working directory for entire Rmd
```{r setup}
knitr::opts_knit$set(root.dir = "../")
```



# Load packages and meta-data (selection of patients with CVID, other PADs and HCs)
# Select correct tube (panel)
```{r}

library(tidyverse)
library(randomForest)
library(readxl)
library(e1071)
library(qsub)
library(ggplot2)
# ------------------------- reading in tube files -----------------------------

allfiles <- list.files("../../PhD/CVID/Excelsheets_data_verwerking", 
                       pattern = ".*_all_data" ,recursive = TRUE, 
                       full.names = TRUE)

# ---------------------------- Class to use -----------------------------------

HC_CVID <- c("HC", "CVID")
PID_HC <- c(HC_CVID, "IgG_deficiency", "IgG_subclass_deficiency", "IgM_deficiency",
            "IgA_deficiency")

group_to_use <- PID_HC  # Choose between HC_CVID and PID_HC
PID <- "_PID" #Set to empty character if you are only working with HC en CVID (otherwise: _PID)

seed <- 1


# --------------------------- Read meta-data ----------------------------------

meta_data_file <- "../../PhD/CVID/Clinical_features_data.txt"
meta_data <- read.delim(meta_data_file, header = TRUE, sep = "\t", 
                        check.names = FALSE) %>% 
  dplyr::filter(diagnosis_group %in% group_to_use) %>%
  dplyr::filter(SAMPLE_ID != "PIDHC081")

if(length(group_to_use) == length(PID_HC)){
  meta_data$diagnosis_group <- as.character(meta_data$diagnosis_group)
  meta_data[!(meta_data$diagnosis_group %in% HC_CVID),"diagnosis_group"] <- "PAD"
  meta_data$diagnosis_group <- as.factor(meta_data$diagnosis_group)
}

# PIDHC081 is removed because he has too few cells for tube 2
rownames(meta_data) <- meta_data$SAMPLE_ID



# ------------------------- read in correct matrix ----------------------------


## Tube is 1 (PBMCs), 2 (B cell subsets) or 5 (T cell subsets) or 1_5 (Combination of all tubes)

tube <- "1_5"

if (tube != "1_5"){
  giant_matrix <- readRDS(file = paste0("Manual_gating/Data/tube",tube,".rds"))
  cols_not_to_convert <- c(1,2,3)
  giant_matrix[,-c(1,2,3)] <- apply(giant_matrix[,-c(1,2,3)], 2, type.convert)
} else {
  giant_matrix <- readRDS(file = "Manual_gating/Data/tube1til5.rds")
  giant_matrix[,-c(1,2,3)] <- apply(giant_matrix[,-c(1,2,3)], 2, type.convert)
  cols_not_to_convert <- c(1,2,3)
}



```


# Random FOrest - NOFS

```{r}
FS <- "NoFS"
tube <- "1_5"



for (tube in c(1,2,5, "1_5")){


if (tube != "1_5"){
  giant_matrix <- readRDS(file = paste0("Manual_gating/Data/tube",tube,".rds"))
  cols_not_to_convert <- c(1,2,3)
  giant_matrix[,-c(1,2,3)] <- apply(giant_matrix[,-c(1,2,3)], 2, type.convert)
} else {
  giant_matrix <- readRDS(file = "Manual_gating/Data/tube1til5.rds")
  giant_matrix[,-c(1,2,3)] <- apply(giant_matrix[,-c(1,2,3)], 2, type.convert)
  cols_not_to_convert <- c(1,2,3)
}

# Leave out next patients because they are missing tube 3  values
giant_matrix <- giant_matrix %>% 
  filter(!(`SAMPLE ID` %in% c("PID057", "PID217", "PIDHC089")))


giant_matrix$Diagnosis_2 <- as.character(giant_matrix$Diagnosis)

giant_matrix$Diagnosis_2[which(giant_matrix$Diagnosis == "PAD")] <- "No_CVID"
giant_matrix$Diagnosis_2[which(giant_matrix$Diagnosis == "HC")] <- "No_CVID"

giant_matrix$Diagnosis_2 <- as.factor(giant_matrix$Diagnosis_2)

giant_matrix <- giant_matrix %>% select(Diagnosis_2, everything())

cols_not_to_convert <- c(1,2,3,4)


meta_data <- meta_data %>% filter(`SAMPLE_ID` %in% giant_matrix$`SAMPLE ID`)
set.seed(seed)

source("R_Scripts/Z_score.R")

  set.seed(1)
  
  train_data <- giant_matrix
  
  
  z_scored_train_data_full <- suppressWarnings(UseZScore(to_be_scored = train_data, 
                                                         cols_not_to_convert,
                                                         meta_data, 
                                                         verbose = FALSE, 
                                                         start = 5, 
                                                         stop = dim(train_data)[2]))
  z_scored_train_data <- z_scored_train_data_full$z_score_matrix
  
  
  # For the three variables
  res3 <- randomForest(x = z_scored_train_data[,-c(1:4)], 
                       y = z_scored_train_data$Diagnosis)
  
  
  # For the two variables
  res2 <- randomForest(x = z_scored_train_data[,-c(1:4)], 
                       y = as.factor(z_scored_train_data$Diagnosis_2))
  
  feat_imp_df_2 <- importance(res2) %>% 
      data.frame() %>% 
      mutate(feature = sub("_POSITIVE.*","",row.names(.))) %>% 
      top_n(20, MeanDecreaseGini)
    
    # plot dataframe
    p_2 <- ggplot(feat_imp_df_2, aes(x = reorder(feature, MeanDecreaseGini), 
                            y = MeanDecreaseGini)) +
      geom_bar(stat='identity') +
      coord_flip() +
      theme_classic() +
      labs(
        x     = "Feature",
        y     = "Importance",
        title = paste0(FS, ": Feature Importance (2 classes), Manual features (MeanDecreaseGini)")) + 
          theme(plot.title = element_text( size = 8) )
    
    feat_imp_df_3 <- importance(res3) %>% 
      data.frame() %>% 
      mutate(feature = sub("_POSITIVE.*","",row.names(.))) %>% 
      top_n(20, MeanDecreaseGini)
    
    # plot dataframe
    p_3 <- ggplot(feat_imp_df_3, aes(x = reorder(feature, MeanDecreaseGini), 
                            y = MeanDecreaseGini)) +
      geom_bar(stat='identity') +
      coord_flip() +
      theme_classic() +
      labs(
        x     = "Feature",
        y     = "Importance",
        title = paste0(FS, ": Feature Importance (3 classes), Manual features (MeanDecreaseGini)")) + 
          theme(plot.title = element_text( size = 8) )
    
    ggsave(
  paste0("Manual_gating/NoFS_2classes_noFStube",tube,".png"),
  plot = p_2,
  width = 10,
  height = 10,
  dpi = 300
)
    
        ggsave(
  paste0("Manual_gating/NoFS_3classes_noFStube",tube,".png"),
  plot = p_3,
  width = 10,
  height = 10,
  dpi = 300
)
    
    plot(p_2)
    plot(p_3)
  

    
saveRDS(res2, file = paste0("Manual_gating/NoFS_res2_tube",tube, ".rds")) 
saveRDS(res3, file = paste0("Manual_gating/No_FS_res3_tube",tube, ".rds"))

}    
    
```


# SVM - NOFS

```{r}  
  
FS <- "NoFS"
tube <- "1_5"



for (tube in c(1,2,5, "1_5")){


if (tube != "1_5"){
  giant_matrix <- readRDS(file = paste0("Manual_gating/Data/tube",tube,".rds"))
  cols_not_to_convert <- c(1,2,3)
  giant_matrix[,-c(1,2,3)] <- apply(giant_matrix[,-c(1,2,3)], 2, type.convert)
} else {
  giant_matrix <- readRDS(file = "Manual_gating/Data/tube1til5.rds")
  giant_matrix[,-c(1,2,3)] <- apply(giant_matrix[,-c(1,2,3)], 2, type.convert)
  cols_not_to_convert <- c(1,2,3)
}



# Leave out next patients because they are missing tube 3  values
giant_matrix <- giant_matrix %>% 
  filter(!(`SAMPLE ID` %in% c("PID057", "PID217", "PIDHC089")))


giant_matrix$Diagnosis_2 <- as.character(giant_matrix$Diagnosis)

giant_matrix$Diagnosis_2[which(giant_matrix$Diagnosis == "PAD")] <- "No_CVID"
giant_matrix$Diagnosis_2[which(giant_matrix$Diagnosis == "HC")] <- "No_CVID"

giant_matrix$Diagnosis_2 <- as.factor(giant_matrix$Diagnosis_2)

giant_matrix <- giant_matrix %>% select(Diagnosis_2, everything())

cols_not_to_convert <- c(1,2,3,4)


meta_data <- meta_data %>% filter(`SAMPLE_ID` %in% giant_matrix$`SAMPLE ID`)
set.seed(seed)


source("R_Scripts/Z_score.R")
set.seed(1)
  
  train_data <- giant_matrix 
  
  
  z_scored_train_data_full <- suppressWarnings(UseZScore(to_be_scored = train_data, 
                                                         cols_not_to_convert,
                                                         meta_data, 
                                                         verbose = FALSE, 
                                                         start = 5, 
                                                         stop = dim(train_data)[2]))
  z_scored_train_data <- z_scored_train_data_full$z_score_matrix
  
  
  
  # For the three classes
  res3 <- svm(Diagnosis ~ ., data = z_scored_train_data[,-c(1,3,4)], kernel = "linear", cost =1, scale = TRUE)
  
  
  # For the two classes
  res2 <- svm(Diagnosis_2 ~ ., data = z_scored_train_data[,-c(2,3,4)], kernel = "linear", cost =1, scale = TRUE)
  
  
  results_2 <- t(res3$coefs) %*% res3$SV 
  results_2 <- apply(abs(results_3), 2, sum)
  
  resultslist_3 <- results_3 %>% abs %>% sort %>% tail(., 3)
 
  
  results_3 <- t(res3$coefs) %*% res3$SV 
  results_3 <- apply(abs(results_3), 2, sum)
  
  resultslist_3 <- results_3 %>% abs %>% sort %>% tail(., 3)
 
  saveRDS(results_2, file = paste0("Manual_gating/NoFS_SVM_res2_tube",tube, ".rds")) 
  saveRDS(results_3, file = paste0("Manual_gating/NoFS_SVM_res3_tube",tube, ".rds"))
 
} 
  

```

# Random FOrest - WithFS
```{r}

FS <- "WithFS"

for (tube in c(1,2,5, "1_5")){


if (tube != "1_5"){
  giant_matrix <- readRDS(file = paste0("Manual_gating/Data/tube",tube,".rds"))
  cols_not_to_convert <- c(1,2,3)
  giant_matrix[,-c(1,2,3)] <- apply(giant_matrix[,-c(1,2,3)], 2, type.convert)
} else {
  giant_matrix <- readRDS(file = "Manual_gating/Data/tube1til5.rds")
  giant_matrix[,-c(1,2,3)] <- apply(giant_matrix[,-c(1,2,3)], 2, type.convert)
  cols_not_to_convert <- c(1,2,3)
}
# Leave out next patients because they are missing tube 3  values
giant_matrix <- giant_matrix %>% 
  filter(!(`SAMPLE ID` %in% c("PID057", "PID217", "PIDHC089")))



giant_matrix$Diagnosis_2 <- as.character(giant_matrix$Diagnosis)

giant_matrix$Diagnosis_2[which(giant_matrix$Diagnosis == "PAD")] <- "No_CVID"
giant_matrix$Diagnosis_2[which(giant_matrix$Diagnosis == "HC")] <- "No_CVID"

giant_matrix$Diagnosis_2 <- as.factor(giant_matrix$Diagnosis_2)

giant_matrix <- giant_matrix %>% select(Diagnosis_2, everything())

cols_not_to_convert <- c(1,2,3,4)


meta_data <- meta_data %>% filter(`SAMPLE_ID` %in% giant_matrix$`SAMPLE ID`)
set.seed(seed)

selected_featureslist_2 <- list()
selected_featureslist_3 <- list()

source("R_Scripts/Z_score.R")

  set.seed(1)
  
  train_data <- giant_matrix
  
  
  z_scored_train_data_full <- suppressWarnings(UseZScore(to_be_scored = train_data, 
                                                         cols_not_to_convert,
                                                         meta_data, 
                                                         verbose = FALSE, 
                                                         start = 5, 
                                                         stop = dim(train_data)[2]))
  z_scored_train_data <- z_scored_train_data_full$z_score_matrix
  
  
  handle_3 <- qsub_lapply(
    X = c(5:(dim(z_scored_train_data)[2])),
    qsub_config = override_qsub_config(
      name = "Kruskal_Wallis_test", 
      verbose = F,
      wait = T,
      memory = "4G"
    ),
    qsub_environment = c("z_scored_train_data"),
    FUN = function(x){
      set.seed(1)
      library(tidyverse)
      
      train_data_new <-  z_scored_train_data %>% select(x,'Diagnosis')
      test <- kruskal.test(rownames(z_scored_train_data[x]) ~ Diagnosis, data = train_data_new )
      
      p_value <- test$p.value
      names(p_value) <- colnames(z_scored_train_data)[x]
      return(p_value)
    } )
  
  # For the three classes
  na_removed_3 <- unlist(handle_3[!is.na(handle_3)])
  adj_p_3 <- stats::p.adjust(na_removed_3, "BH")
  sorted_3 <- sort(adj_p_3)
  z_scored_train_data <- z_scored_train_data %>% select("SAMPLE ID",  "Diagnosis", "Diagnosis_2", names(sorted_3))
  selected_features_3 <- c(4,5)
  for(i_3 in 6:length(sorted_3)){
    max_3 <- 0
    for(f_3 in selected_features_3){
      if(sd(z_scored_train_data %>% .[[f_3]]) > 0 & sd(z_scored_train_data %>% .[[i_3]]) >0 ){
        tmp_3 <- abs(cor(z_scored_train_data %>% .[[i_3]],z_scored_train_data %>% .[[f_3]]))
        if(tmp_3 > max_3) max_3 <- tmp_3
      } else {
        max_3 <- 1
      }
    }
    if(max_3 < 0.2){
      selected_features_3 <- c(selected_features_3,i_3)
      print(selected_features_3)
    } 
  }
  
  
  selected_featureslist_3 <- colnames(z_scored_train_data)[selected_features_3]
  
  selected_train_data_3 <- z_scored_train_data %>% select("SAMPLE ID", "Diagnosis", "Diagnosis_2", selected_featureslist_3)

  
  

  res3 <- randomForest(x = selected_train_data_3[,-c(1:4)], 
                       y = selected_train_data_3$Diagnosis)
  
  
  # For the two classes
  handle_2 <- qsub_lapply(
    X = c(5:dim(z_scored_train_data)[2]),
    qsub_config = override_qsub_config(
      name = "Wilcox_test", 
      verbose = F,
      wait = T,
      memory = "4G"
    ),
    qsub_environment = c("z_scored_train_data"),
    FUN = function(x){
      set.seed(1)
      library(tidyverse)
      test <- wilcox.test(z_scored_train_data %>% filter(Diagnosis_2 == "CVID") %>% 
                            .[[x]],
                          z_scored_train_data %>% filter(Diagnosis_2 == "No_CVID") %>% 
                            .[[x]])
      p_value <- test$p.value
      names(p_value) <- colnames(z_scored_train_data)[x]
      return(p_value)
    } )
  
  na_removed_2 <- unlist(handle_2[!is.na(handle_2)])
  na_removed_2 <- na.omit(na_removed_2)
  adj_p_2 <- stats::p.adjust(na_removed_2, "BH")
  sorted_2 <- sort(adj_p_2)
  z_scored_train_data <- z_scored_train_data %>% select("SAMPLE ID", "Diagnosis", "Diagnosis_2", names(sorted_2))
  selected_features_2 <- c(4,5)
  for(i_2 in 6:length(sorted_2)){
    max_2 <- 0
    for(f_2 in selected_features_2){
      if(sd(z_scored_train_data %>% .[[f_2]]) > 0 & sd(z_scored_train_data %>% .[[i_2]]) >0 ){
        tmp_2 <- abs(cor(z_scored_train_data %>% .[[i_2]],z_scored_train_data %>% .[[f_2]]))
        if(tmp_2 > max_2) max_2 <- tmp_2
      } else {
        max_2 <- 1
      }
    }
    if(max_2 < 0.2){
      selected_features_2 <- c(selected_features_2,i_2)
      print(selected_features_2)
    } 
  }
  
  selected_featureslist_2 <- colnames(z_scored_train_data)[selected_features_2]
  
  selected_train_data_2 <- z_scored_train_data %>% select("SAMPLE ID", "Diagnosis", "Diagnosis_2", selected_featureslist_2)
  
  
  # For the two variables
  res2 <- randomForest(x = selected_train_data_2[,-c(1:4)], 
                       y = as.factor(selected_train_data_2$Diagnosis_2))
  
  
    feat_imp_df_2 <- importance(res2) %>% 
      data.frame() %>% 
      mutate(feature = sub("_POSITIVE.*","",row.names(.))) %>% 
      top_n(20, MeanDecreaseGini)
    
    # plot dataframe
    p_2 <- ggplot(feat_imp_df_2, aes(x = reorder(feature, MeanDecreaseGini), 
                            y = MeanDecreaseGini)) +
      geom_bar(stat='identity') +
      coord_flip() +
      theme_classic() +
      labs(
        x     = "Feature",
        y     = "Importance",
        title = paste0(FS, ": Feature Importance (2 classes), Manual features (MeanDecreaseGini)")) + 
          theme(plot.title = element_text( size = 8) )
    
    feat_imp_df_3 <- importance(res3) %>% 
      data.frame() %>% 
      mutate(feature = sub("_POSITIVE.*","",row.names(.))) %>% 
      top_n(20, MeanDecreaseGini)
    
    # plot dataframe
    p_3 <- ggplot(feat_imp_df_3, aes(x = reorder(feature, MeanDecreaseGini), 
                            y = MeanDecreaseGini)) +
      geom_bar(stat='identity') +
      coord_flip() +
      theme_classic() +
      labs(
        x     = "Feature",
        y     = "Importance",
        title = paste0(FS, ": Feature Importance (3 classes), Manual features (MeanDecreaseGini)")) + 
          theme(plot.title = element_text( size = 8) )
    
    ggsave(
  paste0("Manual_gating/WithFS_2classes_noFStube",tube,".png"),
  plot = p_2,
  width = 10,
  height = 10,
  dpi = 300
)
    
        ggsave(
  paste0("Manual_gating/WithFS_3classes_noFStube",tube,".png"),
  plot = p_3,
  width = 10,
  height = 10,
  dpi = 300
)
    
    plot(p_2)
    plot(p_3)
  


saveRDS(selected_featureslist_2, file = paste0("Manual_gating/WithFS_selected_features_res2_tube",tube, ".rds")) 
saveRDS(selected_featureslist_3, file = paste0("Manual_gating/WithFS_selected_features_res3_tube",tube, ".rds"))
    
saveRDS(res2, file = paste0("Manual_gating/WithFS_res2_tube",tube, ".rds")) 
saveRDS(res3, file = paste0("Manual_gating/WithFS_res3_tube",tube, ".rds"))

}
  

```


# SVM with FS
```{r}



FS <- "WithFS"

for (tube in c(1,2,5, "1_5")){


if (tube != "1_5"){
  giant_matrix <- readRDS(file = paste0("Manual_gating/Data/tube",tube,".rds"))
  cols_not_to_convert <- c(1,2,3)
  giant_matrix[,-c(1,2,3)] <- apply(giant_matrix[,-c(1,2,3)], 2, type.convert)
} else {
  giant_matrix <- readRDS(file = "Manual_gating/Data/tube1til5.rds")
  giant_matrix[,-c(1,2,3)] <- apply(giant_matrix[,-c(1,2,3)], 2, type.convert)
  cols_not_to_convert <- c(1,2,3)
}

# Leave out next patients because they are missing tube 3  values
giant_matrix <- giant_matrix %>% 
  filter(!(`SAMPLE ID` %in% c("PID057", "PID217", "PIDHC089")))


logFile = paste0("Manual_gating/LogFiles/log_file_for_manual_gating_tube",tube,
                 "_WihFS_SVM.txt")
cat(paste0("This is a log file for the patients that were predicted wrong from the RF without a certain experiment day"), file=logFile, append=FALSE, sep = "\n")


giant_matrix$Diagnosis_2 <- as.character(giant_matrix$Diagnosis)

giant_matrix$Diagnosis_2[which(giant_matrix$Diagnosis == "PAD")] <- "No_CVID"
giant_matrix$Diagnosis_2[which(giant_matrix$Diagnosis == "HC")] <- "No_CVID"

giant_matrix$Diagnosis_2 <- as.factor(giant_matrix$Diagnosis_2)

giant_matrix <- giant_matrix %>% select(Diagnosis_2, everything())

cols_not_to_convert <- c(1,2,3,4)


meta_data <- meta_data %>% filter(`SAMPLE_ID` %in% giant_matrix$`SAMPLE ID`)
set.seed(seed)

source("R_Scripts/Z_score.R")

  set.seed(1)
  
  #print(paste0("Experiment day ", experiment))
  train_data <- giant_matrix
  
  
  z_scored_train_data_full <- suppressWarnings(UseZScore(to_be_scored = train_data, 
                                                         cols_not_to_convert,
                                                         meta_data, 
                                                         verbose = FALSE, 
                                                         start = 5, 
                                                         stop = dim(train_data)[2]))
  z_scored_train_data <- z_scored_train_data_full$z_score_matrix
  
  selected_features_list_2 <-  readRDS(file = paste0("Manual_gating/WithFS_selected_features_res2_tube",tube, ".rds")) 
  selected_featurenames_2 <- selected_featureslist_2[[(experiment-12)]]
  
    selected_features_list_3 <-  readRDS(file = paste0("Manual_gating/WithFS_selected_features_res3_tube",tube, ".rds"))
  selected_featurenames_3 <- selected_featureslist_3[[(experiment-12)]]
  
  selected_train_data_2 <- z_scored_train_data %>% select( "Diagnosis_2", "Diagnosis",  "SAMPLE ID", "EXPERIMENT NAME", selected_featurenames_2)
  
  selected_train_data_3 <- z_scored_train_data %>% select("Diagnosis_2", "Diagnosis",  "SAMPLE ID", "EXPERIMENT NAME", selected_featurenames_3)

  
  
  # For the three classes
  res3 <- svm(Diagnosis ~ ., data = selected_train_data_3[,-c(1,3,4)], kernel = "linear", cost =1, scale = TRUE)

  # For the two classes
  res2 <- svm(Diagnosis_2 ~ ., data = selected_train_data_2[,-c(2,3,4)], kernel = "linear", cost =1, scale = TRUE)
  
  results_3 <- t(res3$coefs) %*% res3$SV 
  results_3 <- apply(abs(results_3), 2, sum)
  
  resultslist_3 <- results_3 %>% abs %>% sort %>% tail(., 3)
  
  saveRDS(results_2, file = paste0("Manual_gating/NoFS_SVM_res2_tube",tube, ".rds")) 
  saveRDS(results_3, file = paste0("Manual_gating/NoFS_SVM_res3_tube",tube, ".rds"))
 
  
}  

```

# Make tsne plots
```{r}

set.seed(seed)

library(Rtsne)
library(gridExtra)
source("R_Scripts/Z_score.R")


meta_data_file <- "../../PhD/CVID/Clinical_features_data.txt"
meta_data <- read.delim(meta_data_file, header = TRUE, sep = "\t", 
                        check.names = FALSE) %>% 
  dplyr::filter(diagnosis_group %in% group_to_use) %>%
  dplyr::filter(SAMPLE_ID != "PIDHC081")

if(length(group_to_use) == length(PID_HC)){
  meta_data$diagnosis_group <- as.character(meta_data$diagnosis_group)
  meta_data[!(meta_data$diagnosis_group %in% HC_CVID),"diagnosis_group"] <- "PAD"
  meta_data$diagnosis_group <- as.factor(meta_data$diagnosis_group)
}

# PIDHC081 is removed because he has too few cells for tube 2
rownames(meta_data) <- meta_data$SAMPLE_ID
meta_data <- meta_data %>% mutate("diagnosis_group_2" = ifelse(meta_data$diagnosis_group == "CVID", "CVID", "No_CVID") )

annotation <- meta_data


cols_not_to_convert <- c(1:3)
  
  z_scored_train_data_full <- suppressWarnings(UseZScore(to_be_scored = giant_matrix, 
                                                         cols_not_to_convert,
                                                         meta_data, 
                                                         verbose = FALSE, 
                                                         start = 4, 
                                                         stop = ncol(giant_matrix)))
  z_scored_train_data <- z_scored_train_data_full$z_score_matrix
  


## Features were obtained from SVM model of the three panels seperately without FS  
  
z_scored_train_data <- z_scored_train_data %>% select("SAMPLE ID", "Diagnosis", "EXPERIMENT NAME", "B cells (% lymphos)", "HLADR relative MFI (B cells)", "CD4+ T cells (% lymphos)", "IgG-IgA+ (% sw mem B cells)", "CD138+ plasma cells (% alive)", "IgD-CD27- B cells (% B cells)", "RO-CCR7+ (% CD8+ T cells)                          naive", "CXCR5 / CCR7 relative MFI (B cells) ratio", "naive CD8/ total memory CD8 ratio")
  


  
  tsne_res <- Rtsne(z_scored_train_data[,-c(1:3)], perplexity = 15)

annotation <- annotation %>% 
  mutate(tsne_1 = tsne_res$Y[,1]) %>% 
  mutate(tsne_2 = tsne_res$Y[,2])

p <- annotation %>% 
  ggplot(aes(x = tsne_1, y = tsne_2)) + 
  geom_point(size = 5) +
  theme_minimal() + 
  ggtitle("A. TSNE based on Manual features")

p_gender <- annotation %>% 
  ggplot(aes(x = tsne_1, y = tsne_2)) + 
  geom_point(aes(color = gender), size = 5) + guides(colour = guide_legend( title = "Gender")) +
  theme_minimal() + 
  ggtitle("B.")

  
p_diagnosis <- annotation %>% 
  ggplot(aes(x = tsne_1, y = tsne_2)) + 
  geom_point(aes(color = diagnosis_group), size = 5) +  guides(colour = guide_legend( title = "Diagnosis")) +
  theme_minimal() + 
  ggtitle("C.")


p_age <- annotation %>% 
  ggplot(aes(x = tsne_1, y = tsne_2)) + 
  geom_point(aes(color = as.numeric(gsub(",", ".", gsub("\\.", "", age_y)))), size = 5) +guides(color=guide_legend(title="Age"))
p_age <- p_age +  theme_minimal()  +
  ggtitle("D.")

 
l <- list(p,p_gender, p_diagnosis, p_age)

ggsave("Manual_gating/tsne_new2.png", arrangeGrob(grobs = l), width = 10, height = 10, dpi = 300)

plot(p_age)




p_diagnosis <- annotation %>% 
  ggplot(aes(x = tsne_1, y = tsne_2)) + 
  geom_point(aes(color = diagnosis_group), size = 5) +  guides(colour = guide_legend( title = "Diagnosis")) +
  theme_minimal() + geom_text(aes(label=SAMPLE_ID),hjust=0, vjust=2)


p_diagnosis

```

