---
title: "New_quality_control"
output: html_document
editor_options: 
  chunk_output_type: inline
---

## Set working directory for entire Rmd
```{r setup}
knitr::opts_knit$set(root.dir = "../")
```


## Load libraries and script for quality control functions
```{r}
library(flowCore) # For basic flow cytometry functionality
library(flowWorkspace) # For loading FlowJo X workspaces
library(FlowSOM) # For making FlowSOM visualisations
library(flowAI)
library(tidyverse)
library(pheatmap)
library(RColorBrewer)
library(FlowSOMworkshop)

options("scipen"=100)

source("R_Scripts/Quality_control_functions.R")
```


## Initialise the experiment days that have to be analysed
```{r}
# Give experiment days
start <- 13
end <- 33
i <- 30

tube <- 5

if (tube ==1){ #PBMCs
  cell_types <- c("B cells", "CD4+ T cells", "CD4- T cells", "Monocytes", "basophils", "ILCs", "NK cells", "NKT cells", "DCs", "cDCs", "pDCs")
  start_gate <- c("alive")
  } else if (tube ==2){ #B cells
    cell_types <- c("marginal zone", "naive", "sw memory", "IgD-CD27-", "CD138+ plasma cells", "CD38+ plasmablasts", "CD21low B cells", "CD21-CD38+ B cells", "CD24++CD38++ Tr B cells" )
  start_gate <- c("B cells", "CD38+ plasmablasts")
  } else if (tube == 5){ #T cells
    cell_types <- c( "CD4+ T/Q1: CD45RO#PE-Cy5- , CCR7#PE+", "CD4+ T/Q2: CD45RO#PE-Cy5+ , CCR7#PE+" , "CD4+ T/Q3: CD45RO#PE-Cy5+ , CCR7#PE-",  "CD4+ T/Q4: CD45RO#PE-Cy5- , CCR7#PE-", "CD31+RO-CD4+", "CD8+ T/Q1: CD45RO#PE-Cy5- , CCR7#PE+",  "CD8+ T/Q2: CD45RO#PE-Cy5+ , CCR7#PE+", "CD8+ T/Q3: CD45RO#PE-Cy5+ , CCR7#PE-", "CD8+ T/Q4: CD45RO#PE-Cy5- , CCR7#PE-", "CD31+RO-CD8+", "gd T cells", "Treg")
    start_gate <- c("T cells")
  }
```

# Preprocessing and quality control of the data
```{r}

i <- 14

# Initialising the different directories 
#dir <- "/media/sf_Fedora_General_Folder/PhD/CVID/CVID_Delfien_orig"
dir <- "/Fedora_General_Folder/PhD/CVID/CVID_Delfien_orig"
dataDir <- paste0("Preprocessing/Tube",tube)
suppressWarnings(dir.create(dataDir))


for (i in start:end){
  
  print(paste0("Working on experiment ", i))
 

  fcsDirs <- file.path(dir,paste0("Experiment CVID_",i))
  if (tube != 3){ daysDirs <- 
    list.files(fcsDirs, pattern = "^CVID_...D1.*[0-9]$")} else {
      daysDirs <- list.files(fcsDirs, pattern = "^CVID_...D3.*[0-9]$")
    }
  fcsDir <- file.path(fcsDirs, daysDirs)
  


  if (tube %in% c(1,2,6)){
  files <- list.files(file.path(fcsDir, "Tube 1-2-6"), pattern = paste0("^P.*Tube ",tube), ignore.case = TRUE)
  if (length(files) ==0) {files <- 
    list.files(file.path(fcsDir, "CVID plate 1-2-6"), pattern = paste0("^P.*Tube ",tube), ignore.case = TRUE)}
  } else if (tube == 5){
     files <- list.files(file.path(fcsDir, "Tube 5"),pattern = paste0("^P.*Tube ",tube), ignore.case = TRUE)
  if (length(files) ==0) {files <- 
    list.files(file.path(fcsDir, "CVID plate 5"), pattern = paste0("^P.*Tube ",tube), ignore.case = TRUE)}
  }
  
  # Workspace file
  wsp_file <- file.path(dir, paste0("Experiment CVID_",i,"/Exp CVID_",i," tube ",tube,".wsp"))
  # Create some directories:
  #       To save the QC figures
  #       To save the preprocessed fcs files
  #       To save the FlowSOM results
  
  qcDir <- file.path(dataDir, paste0("Exp_",i,"/QC"))
  filesDir <- file.path(dataDir, paste0("Exp_",i))
  
  suppressWarnings(dir.create(file.path(dataDir, paste0("Exp_",i))))
  suppressWarnings(dir.create(qcDir))
  
  ###################
  ## Preprocessing ##
  ###################
  
  
  tubes <- parse_flowjo_compensation(files, 
                                     wsp_file, 
                                     fcsDir, 
                                     tube, 
                                     experiment = i, 
                                     plot = FALSE, 
                                     html_plot = html_plot) # set HTML_plot to "_QC" or FALSE
  
  if (tube == 2){
    gating_matrix <- tubes$gates
    saveRDS(gating_matrix, file = paste0("Preprocessing/Tube2/Gatingmatrix_exp",i,".Rds"))
  }
  
  data_live <- gating_subset_adapted(tubes, start_gate)
  #plot_aggregate(flowset = data_live$flowSet, output_image = paste0("Preprocessing/Tube",tube,"/Exp_",i,"/QC/aggregate",i,".png"))
  
  item <- 1
  for (item in 1:length(data_live$flowSet)) { 
    
    
    print(paste0("Saving file ", data_live$flowSet[[item]]@description$GUID))
    suppressWarnings( write.FCS(data_live$flowSet[[item]],
              file = file.path(filesDir,
                               gsub(".fcs","_pp.fcs",data_live$flowSet[[item]]@description$GUID))))
    
    if (tube == 3){
      manual_labels <- manual_vector_tube3(data_live$gates[[item]])
     
    } else {manual_labels <- manual_vector(data_live$gates[[item]], cell_types)}
    
    manual <- list("matrix" = data_live$gates[[item]],
                   "manual" = manual_labels)
    
    save(manual,
         file = file.path(filesDir,
                          gsub(".fcs","_manual.Rdata",data_live$flowSet[[item]]@description$GUID)))
  }
}

```


# make factor vector of manual labels if an error occured in building FlowSOM models
```{r}

files <- list.files(file.path(filesDir), pattern = ".*.Rdata", recursive = TRUE, full.names = TRUE)
file <- files[1]





for (file in files){
  load(file)
 new_manual_labels <- as.character(manual$manual)
 test <- sub("(.*)#.*([+-] .*)#.*([+-])", "\\1\\2\\3", new_manual_labels)
  
  
  manual_labels <- as.factor(test)
  manual$manual <- manual_labels
  
  save(manual, file = file)
  
  
}

```


## Make an overview figure of FlowAI quality control
# Make sure that there are no duplicate rowfiles in qcmini.txt
```{r}
tube <- 5
start <- 13
end <- 33

breaksList =  seq(0, 100, by = 1)

i <- 13

pdf(file = paste0("Preprocessing/Tube",tube,"/QC_overview.pdf"))

for (i in start:end){
  print(i)
  qc_mini <- paste0("Preprocessing/Tube",tube,"/Exp_",i,"/QC/QCmini.txt")
  pctg <- read.delim(qc_mini, check.names = FALSE) %>% 
    column_to_rownames("Name file") %>% 
    select(matches('%|n. of events'))
  


annotation <- data.frame(nCells = pctg$`n. of events`, row.names = rownames(pctg))
  pheatmap(pctg[,-1],
           cluster_rows = FALSE,
           cluster_cols = FALSE,
           main = paste0("Experiment ", i), 
           breaks = breaksList,
           color= colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(length(breaksList)),
           annotation_row = annotation,
           annotation_colors = list(nCells= c("white", "purple"))
  )
}

dev.off()
```


